### 2023-03-13 main/chapter_2ljets.tex
add uncertainties summary plot and commentary

Throughout $\twoljets$-electroweak, simulated Monte Carlo samples are the
primary estimates of background and signal processes.
We first simulate samples at `truth-level' by random sampling from
distributions with parton distribution functions and scattering matrix elements
that are implemented by various software packages to approximate theoretical
predictions from quantum field theory.
Simulating events from high-order perturbative quantum field theory is
difficult and computationally expensive, so these implementations use
lower-order approximations, which we normalize to cross-section predictions
that are evaluated at higher precision.
Parton shower programs are used to model the soft QCD evolution of
strongly-interacting particles from the hard scatter down to the low energy
scales of hadronization; because of colour confinement, only these
(colour-neutral) hadrons can survive to interact with the detector.
The various Monte Carlo simulation programs used in this search are summarized
in Table~\ref{tab:2ljets_paper_mc}, which is adapted from the $\twoljets$
paper.
Truth-level results are finally passed through a full simulation of the
\atlas\ detector to simulate `reconstruction-level'
data~\cite{SOFT-2010-01, geant2003}.

@@

\begin{sidewaystable}[tp]
\begin{center}
\scriptsize
\begin{tabular}{llllll}
Physical process
    & Generator
    & Parton Shower
    & Cross-section order in QCD
    & Tune
    & PDF set
\\[0.2em]
\hline
\\[0.2em]
$\ttbar$~\cite{ATL-PHYS-PUB-2016-004}
    & \powhegbox\,v2~\cite{Nason:2004rx,Frixione:2007vw,Alioli:2010xd}
    & \pythia\,8.230~\cite{Sjostrand:2014zea}
    & NNLO+NNLL~\cite{ttbarxsec1,ttbarxsec2,ttbarxsec3,ttbarxsec4,ttbarxsec5}
    & A14~\cite{ATL-PHYS-PUB-2014-021}
    & \textsc{NNPDF2.3lo}~\cite{Ball:2012cx}
\\
$Wt$~\cite{ATL-PHYS-PUB-2016-004}
    & \powhegbox\,v2
    & \pythia\,8.230
    & NNLO+NNLL~\cite{Kidonakis:2010b,Kidonakis:2010tc,Kidonakis:2011wy,Frederix:2012dh}
    & A14
    & \textsc{NNPDF3.0nlo}~\cite{Ball:2014uwa}
\\
$VV$~\cite{ATL-PHYS-PUB-2016-002}
    & \sherpa\,2.2.2~\cite{Bothmann:2019yzt}
    & \sherpa\,2.2.2 & NLO~\cite{diboson1,diboson2}
    & \textsc{Sherpa}~default
    & \textsc{NNPDF3.0nnlo}
\\
$Z/\gamma^{*}(\rightarrow \ell \ell)$\,+\,jets~\cite{ATL-PHYS-PUB-2016-003}
    & \sherpa\,2.2.1
    & \sherpa\,2.2.1
    & NNLO~\cite{DYNNLO1,DYcrosssection}
    & \textsc{Sherpa}~default
    &\textsc{NNPDF3.0nnlo}
\\
Higgs ($\ttbar h$)
    & \powhegbox\,v2
    & \pythia\,8.230
    & NLO~\cite{deFlorian:2016spz}
    & A14
    & \textsc{NPDF2.3lo}
\\
Higgs (VBF, $Vh$)
    & \powhegbox\,v2
    & \pythia\,8.212
    & NNLO~\cite{deFlorian:2016spz}
    & AZNLO~\cite{STDM-2012-23}
    & \textsc{CTEQ6L1}~\cite{Pumplin:2002vw}
\\
Higgs (ggF)
    & \powhegbox\,v2
    & \pythia\,8.212
    & NNNLO~\cite{deFlorian:2016spz}
    & AZNLO
    & \textsc{CTEQ6L1}
\\
$ttt$, $tttt$, $\ttbar WW$
    & \amcatnlo~\cite{Alwall:2014hca}
    & \pythia\,8.186~\cite{Skands:2014pea}
    & NLO~\cite{Frederix:2017wme,Alwall:2014hca}
    & A14
    & \textsc{NNPDF2.3lo}
\\
$\ttbar V$~\cite{ATL-PHYS-PUB-2016-005,Garzelli:2012bn}
    & \amcatnlo
    & \pythia\,8.210
    & NLO~\cite{Campbell:2012,Lazopoulos:2008,Alwall:2014hca}
    & \textsc{A14}
    & \textsc{NNPDF2.3lo}
\\
$VVV$~\cite{ATL-PHYS-PUB-2016-002}
    & \sherpa\,2.2.2
    & \sherpa\,2.2.2
    &  NLO~\cite{Bothmann:2019yzt}
    & \textsc{Sherpa}~default
    & \textsc{NNPDF3.0nnlo}
\\
\end{tabular}
\caption{%
Summary of the simulation tools and configurations used for central background
samples in $\twoljets$.
Each row specifies a combination of matrix element generator, parton shower,
and Parton Distribution Function (PDF) components, whose combination we use to
simulate the named physical process.
The tune specifies many parameters of the simulation tools to achieve a good
description of some historical calibration data.
Total sample cross-sections are normalized to a reference value evaluated
at the specified order in perturbative QCD (pQCD).
Symbols:
$V$ is $W$ or $Z$,
VBF is Vector Boson Fusion,
ggF is gluon-gluon Fusion.
This table is adapted from~\cite{atlas2022searches}.
}
\label{tab:2ljets_paper_mc}
\end{center}
\end{sidewaystable}

@@

>>>>

Throughout $\twoljets$-electroweak, simulated Monte Carlo samples are the
primary estimates of background and signal processes.
We first simulate samples at `truth-level' by random sampling from
distributions with parton distribution functions and scattering matrix elements
that are implemented by various software packages to approximate theoretical
predictions from quantum field theory.
Simulating events from high-order perturbative quantum field theory is
difficult and computationally expensive, so these implementations use
lower-order approximations, which we normalize to cross-section predictions
that are evaluated at higher precision.
Parton shower programs are used to model the soft QCD evolution of
strongly-interacting particles from the hard scatter down to the low energy
scales of hadronization; because of colour confinement, only these
(colour-neutral) hadrons can survive to interact with the detector.
The various Monte Carlo simulation programs used in this search are summarized
in Table~\ref{tab:2ljets_paper_mc}, which is adapted from the $\twoljets$
paper~\cite{atlas2022searches}.
Truth-level results are finally passed through a full simulation of the
\atlas\ detector to simulate `reconstruction-level'
data~\cite{SOFT-2010-01, geant2003}.

@@

\begin{sidewaystable}[tp]
\begin{center}
\scriptsize
\begin{tabular}{llllll}
Physical process
    & Generator
    & Parton Shower
    & Cross-section order in QCD
    & Tune
    & PDF set
\\[0.2em]
\hline
\\[0.2em]
$\ttbar$~\cite{ATL-PHYS-PUB-2016-004}
    & \powhegbox\,v2~\cite{Nason:2004rx,Frixione:2007vw,Alioli:2010xd}
    & \pythia\,8.230~\cite{Sjostrand:2014zea}
    & NNLO+NNLL~\cite{ttbarxsec1,ttbarxsec2,ttbarxsec3,ttbarxsec4,ttbarxsec5}
    & A14~\cite{ATL-PHYS-PUB-2014-021}
    & \textsc{NNPDF2.3lo}~\cite{Ball:2012cx}
\\
$Wt$~\cite{ATL-PHYS-PUB-2016-004}
    & \powhegbox\,v2
    & \pythia\,8.230
    & NNLO+NNLL~\cite{Kidonakis:2010b,Kidonakis:2010tc,Kidonakis:2011wy,Frederix:2012dh}
    & A14
    & \textsc{NNPDF3.0nlo}~\cite{Ball:2014uwa}
\\
$VV$~\cite{ATL-PHYS-PUB-2016-002}
    & \sherpa\,2.2.2~\cite{Bothmann:2019yzt}
    & \sherpa\,2.2.2 & NLO~\cite{diboson1,diboson2}
    & \textsc{Sherpa}~default
    & \textsc{NNPDF3.0nnlo}
\\
$Z/\gamma^{*}(\rightarrow \ell \ell)$\,+\,jets~\cite{ATL-PHYS-PUB-2016-003}
    & \sherpa\,2.2.1
    & \sherpa\,2.2.1
    & NNLO~\cite{DYNNLO1,DYcrosssection}
    & \textsc{Sherpa}~default
    &\textsc{NNPDF3.0nnlo}
\\
Higgs ($\ttbar h$)
    & \powhegbox\,v2
    & \pythia\,8.230
    & NLO~\cite{deFlorian:2016spz}
    & A14
    & \textsc{NPDF2.3lo}
\\
Higgs (VBF, $Vh$)
    & \powhegbox\,v2
    & \pythia\,8.212
    & NNLO~\cite{deFlorian:2016spz}
    & AZNLO~\cite{STDM-2012-23}
    & \textsc{CTEQ6L1}~\cite{Pumplin:2002vw}
\\
Higgs (ggF)
    & \powhegbox\,v2
    & \pythia\,8.212
    & NNNLO~\cite{deFlorian:2016spz}
    & AZNLO
    & \textsc{CTEQ6L1}
\\
$ttt$, $tttt$, $\ttbar WW$
    & \amcatnlo~\cite{Alwall:2014hca}
    & \pythia\,8.186~\cite{Skands:2014pea}
    & NLO~\cite{Frederix:2017wme,Alwall:2014hca}
    & A14
    & \textsc{NNPDF2.3lo}
\\
$\ttbar V$~\cite{ATL-PHYS-PUB-2016-005,Garzelli:2012bn}
    & \amcatnlo
    & \pythia\,8.210
    & NLO~\cite{Campbell:2012,Lazopoulos:2008,Alwall:2014hca}
    & \textsc{A14}
    & \textsc{NNPDF2.3lo}
\\
$VVV$~\cite{ATL-PHYS-PUB-2016-002}
    & \sherpa\,2.2.2
    & \sherpa\,2.2.2
    &  NLO~\cite{Bothmann:2019yzt}
    & \textsc{Sherpa}~default
    & \textsc{NNPDF3.0nnlo}
\\
\end{tabular}
\caption[
Summary of the simulation tools and configurations
]{%
Summary of the simulation tools and configurations used for central background
samples in $\twoljets$.
Each row specifies a combination of matrix element generator, parton shower,
and Parton Distribution Function (PDF) components, whose combination we use to
simulate the named physical process.
The tune specifies many parameters of the simulation tools to achieve a good
description of some historical calibration data.
Total sample cross-sections are normalized to a reference value evaluated
at the specified order in perturbative QCD (pQCD).
Symbols:
$V$ is $W$ or $Z$,
VBF is Vector Boson Fusion,
ggF is gluon-gluon Fusion.
This table is adapted from the $\twoljets$ paper~\cite{atlas2022searches}.
}
\label{tab:2ljets_paper_mc}
\end{center}
\end{sidewaystable}

@@

These parameters control the yield assigned in each region by our
\heplikelihood\ model as described in
Section~\ref{sec:searches_heplikelihood}.
Uncertainties in the parameters therefore propagate to uncertainties in the
background expectations.
To illustrate which categories of parameters contribute the most uncertainty
to which regions,
Figure~\ref{fig:2ljets_fit_uncertainty} presents a summary of the relative
uncertainties attributable to sources of uncertainty in each of region of the
$\twoljets$-electroweak search.
This figure is reproduced from the $\twoljets$ paper~\cite{atlas2022searches}.
No single source of uncertainty dominates in all regions.
Theoretical uncertainties are large in many regions, most dramatically
in SR-High-1J and SR-$\llbb$. These two regions have small yields and large
variations due to alternative sample comparisons and uncertainties that are
not constrained in control regions;
see Section~\ref{sec:2ljets_jet_mass} for details of the extra uncertainties
in SR-High-1J.
Large relative contributions from FNP backgrounds drive the uncertainties in
SR-High-16, both directly in the data-driven uncertainty category and
indirectly through their poor statistical precision in the statistical
uncertainty category.
Experimental and normalization uncertainties make middling contributions
in all regions.

\begin{figure}[tp]
\centering
\adjincludegraphics[width=0.98\textwidth]{figures/2ljets_syst_summary_plot_ewk}
\caption[
Summary of the contributions to uncertainty on total backgrounds
]{%
Summary of the contributions to uncertainty on total backgrounds, represented
as relative errors for all signal regions of the $\twoljets$-electroweak
search.
Uncertainty attributions are assigned by a standard method that explores the
variations in the \heplikelihood\ model with respect to individual parameters.
Experimental uncertainties (``Exp. Unc.'') include detector effects in object
identification and reconstruction, pile-up, and triggering.
Statistical uncertainties (``Bkg. Sample Stats'') arise from the
finite number of weighted Monte Carlo events in our background samples
and from statistical uncertainties on the Matrix Method FNP estimate.
Data-driven background uncertainties (``DD Bkg. Unc.'') include only effects
from the FNP background estimate in this search.
Theoretical uncertainties (``Theory Unc.'') include only theoretical
uncertainties on the Standard Model backgrounds, and not on any signals.
Background normalization uncertainties (``Bkg. Norm.'') derive from the
free normalization factors that should be constrained by control regions.
The total uncertainty (``Total Unc.'') is a sum in quadrature of all
contributions.
\\[0.4em]
This figure is reproduced from the $\twoljets$ paper~\cite{atlas2022searches}.
}
\label{fig:2ljets_fit_uncertainty}
\end{figure}

### 2023-03-13 main/chapter_searches.tex
add uncertainties summary plot and commentary

\subsection{Likelihood modelling}
To represent the Standard Model and alternatives, LHC data analyses use
parametric models to describe the models' Poisson expectations of
collision events and how those expectations change in response to their
parameters.
Parameters often correspond to interpretable, physical quantities and when
those parameters are known only with some uncertainty, the models help to
propagate uncertainty into variations of their outputs.

>>>>

\subsection{Likelihood modelling}
\label{sec:searches_heplikelihood}
To represent the Standard Model and alternatives, LHC data analyses use
parametric models to describe the models' Poisson expectations of
collision events and how those expectations change in response to their
parameters.
Parameters often correspond to interpretable, physical quantities and when
those parameters are known only with some uncertainty, the models help to
propagate uncertainty into variations of their outputs.
